Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-12-09 patchy-review-responses.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We thank the editor and the reviewers for their helpful comment. We
have substantially revised our paper in response ot these comments,
and think that it is substantially improved as a consequence. Below we
have a detailed response point-by-point to the reviewers comments. We
first address the consensus points highlighted by the editor.

@@

We have rewritten the method section to simplify the presentation. 
Before writing the paper, we had initially explored a number of simpler ways of developing the
argument (along similar the lines to those suggested by the reviewers), 
and went around in circles for quite some time
before settling on the argument given in the first draft.
As suggested by the third reviewer, 
we have moved the most difficult part of the argument to the Materials \& Methods,
replacing it in the Results with a more \textit{ad hoc} argument
along the lines of that suggested by the reviewers.
We do think the more formal argument is more solid, however,
and therefore necessary to keep; 
as discussed below in our response to the reviewers, various seemingly reasonable arguments
give different answers for subtle reasons.

Other changes of note are:
\begin{itemize}

  \item Several reviewers pointed out that in approximating $\E[1-(1-p_e)^K] \approx \E[K] p_e$
    we were neglecting the case that $s_m \ll s_p$.
    This omission was sloppy on our part,
    which we have rectified; see the new equation~\eqref{eqn:twocase_migrate}.

  \item We have also improved our discussion of the limits to the theory,
    mostly in the (short) section \nameref{ss:assumptions}.
    The main point of note was that, indeed as Reviewer 3 pointed out,
    we were \gc{originally} running \gc{simulations} in regions where $s_m$ was too small for the theory to apply,
    resulting in $\migrate$ that was not monotonic with $s_m$.
    Happily, the natural restriction of $R \sqrt{s_m}/\sigma > 1$ coincides with the condition
    for $\migrate$ to decrease with $s_m$;
    we don't have good intuition for why the formula should increase in the region where it is inapplicable,
    but this seems beside the point.

@@

\reply
This is a very interesting point,
and the basic tool of a diffusion approximation seems to be the same in both.
However, something is not matching up,
because \citet{barton1979geneflow} finds that the length of the region of differentiated genome 
decays with $1/\sqrt{T}$,
whereas with our approximation the length of the differentiated haplotype decays as $1/T$.
The difference may be that Barton considers LD and we consider a haplotype unbroken by recombination.
There are also other assumptions in Barton's paper that it's applicability unclear.
On the balance, and considering the other reviewers' comments,
we've cut this section and discussed it a bit in the discussion. 
\llname{rr:haplotype_discussion}

%%%%%%%%
\point{In “Applications”,}{the authors spend a lot of time introducing the mice
    and calculating the values for Fig. 6, but then they don’t show where the
    actual data lie on those plots. Is there enough information to do some
    kind of comparison? How close are the nearest samples to each other?
    Have no examples of alleles spreading by migration been found? If not,
    how much of a bound does this put on the parameter values? Is there any
    data on the genetic map length of the region of reduced diversity around
    the selected locus?
}

\reply
At the moment it is only known that the same MC1R alleles do not
contribute to color polymorphism on a number of other patches. We know
from personal discussion with Michael Nachman that another color
polymorphism allele, in a different gene, has been mapped on another
patch. We are currently talking to Nachman about a broader-scale
effort to identify other causal alleles motivated in part by the work described
in the paper. However, this work is too preliminary to offer any thing
certain.

 There is no data about the genetic map in this species, but
as a rough rule of thumb 1cM is roughly 1Mb in a number of mammalian
species, we have added this guide to the physical scale of the shared haplotype in the figure caption.

### 2014-12-09 patchy-selection-paper.tex
small changes throughout response to reviews, and discussion

\begin{figure}[ht]
  \begin{center}
    \includegraphics{Lava_flow_mice_prob_parallel}
  \end{center}
  \caption{
Top panel: The probability of parallel mutation as a function of the distance
between environmental patches for two different cline widths
and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
see text. Bottom panel: The initial genetic length of the haplotype
shared between patches due to adaptation via migration as a function
of the distance between environmental patches. As a rough rule of
thumb $1$cM is approximately $1$Mb in a number of mammalian species. \linelabel{rr:cMMb}
  \label{fig:mice_prob_parallel}
  }
\end{figure}

@@

\paragraph{Decay of shared haplotypes} \nameref{rr:haplotype_discussion}
To provide context for the results on shared haplotype length in \nameref{ss:haplotype_length}
it is important to also understand
the process by which haplotypes are whittled down within patches.
The initial haplotype that sweeps within a patch 
will be dispersed over time by recombination.
Likewise, the haplotype that is shared between patches coadapted by migration
(equation~\ref{eqn:haplotype_length}) will also break down. 
However, a long time after the initial sweep,
we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
than with individuals elsewhere,
since selection against migrants 
decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
The literature on clines \citep[e.g.]{barton1979geneflow} has important information,
but a good deal more work is needed
to provide robust estimates for these processes.

\paragraph{Standing variation} 
We have focused on relative rates of adaptation,
since we are thinking of applications in which we know that adaptation has occurred,
and the question is whether or not adaptations in distinct patches
have appeared independently or not.
However, especially absolute if rates of adaptation by new mutation are very low,
any adaptation that does occur may have to make use of standing variation.
The case of a panmictic population was studied by \citet{softsweepsI},
and we study the case of a continuous, spatial population in an upcoming paper.
If parallelism in local adaptation of the sort we study here is due to standing variation
rather than new mutation,
then the dynamics of adaptation should not depend strongly on migration patterns
(but note that the initial spatial distribution of standing variation may).

>>>>

\begin{figure}[ht]
  \begin{center}
    \includegraphics{Lava_flow_mice_prob_parallel}
  \end{center}
  \caption{
Top panel: The probability of parallel mutation as a function of the distance
between environmental patches for two different cline widths
and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
see text. Bottom panel: The initial genetic length of the haplotype
shared between patches due to adaptation via migration as a function
of the distance between environmental patches. As a rough rule of
thumb $1$cM is approximately $1$Mb in a number of mammalian species. 
  \label{fig:mice_prob_parallel}
  }
\end{figure}

@@

\paragraph{Decay of shared haplotypes} \nameref{rr:haplotype_discussion}
To provide context for the results on shared haplotype length in \nameref{ss:haplotype_length}
it is important to also understand
the process by which haplotypes are whittled down within patches.
The initial haplotype that sweeps within a patch 
will be dispersed over time by recombination.
Likewise, the haplotype that is shared between patches coadapted by migration
(equation~\ref{eqn:haplotype_length}) will also break down. 
However, a long time after the initial sweep,
we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
than with individuals elsewhere,
since selection against migrants 
decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
The literature on clines \citep[e.g.]{barton1979geneflow} has important information,
but more work is needed
to provide robust estimates for these processes.
\gc{We note that questions about the genomic length-scale of signals of
sweeps shared by migration have also been addressed in discrete
population settings \citep{REFSCOMMENTEDBELOW}. This work has shown
that the length of shared sweept haplotype is often significantly
shorter than the sweep with each patch, resulting in a pattern of
shoulders ofelevated $F_{ST}$ between adapted populations some distance away from
the shared selected allele. It would be of interest to see how similar
patterns can arise in a continuous population setting, as a way of
uniting these results. }

%Slatkin and Wiehe and this paper on spatial hitchhiking http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3176130/
\paragraph{Standing variation} 
We have focused on relative rates of adaptation,
since we are thinking of applications in which we know that adaptation has occurred,
and the question is whether or not adaptations in distinct patches
have appeared independently or not.
However, especially absolute if rates of adaptation by new mutation are very low,
any adaptation that does occur may have to make use of standing variation.
The case of a panmictic population was studied by \citet{softsweepsI},
and we study the case of a continuous, spatial population in an upcoming paper.
If parallelism in local adaptation of the sort we study here is due to standing variation
rather than new mutation,
then the dynamics of adaptation should not depend strongly on migration patterns
(but note that the initial spatial distribution of standing variation may).
